---
title: "ML-Tools-Assignment-2"
author: "Gyongyver Kamenar (2103380)"
date: "3/27/2022"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: no
    toc: no
    extra_dependencies: ["float"]
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r}
library(tidyverse)
library(kableExtra)
library(keras)
library(ggplot2)
library(imager)

# Set my theme
devtools::source_url('https://raw.githubusercontent.com/gyongyver-droid/ceu-data-analysis/master/Assignment1/theme_gyongyver.R')
#theme_set(theme_gyongyver())
```


# Problem 1


## A)
**What would be an appropriate metric to evaluate your models? Why?**

Accuracy would be an appropriate metric

## B)
**Get the data and show some example images from the data.**


```{r Get the data}
myseed<-20220403
set.seed(myseed)
data <-dataset_mnist()

train_x<-data$train$x
train_y<-data$train$y

test_x<-data$test$x
test_y<-data$test$y

# Merge features with labels for the plot
train <-cbind(as.tibble(data$train$x),y=data$train$y)

# Factorizing and scaling for plotting
train <- train %>% mutate(
  y = as.factor(y),
  across(-y, ~./255)
)



# Plot the images
rotate <- function(x) t(apply(x, 2, rev))
par(mfrow=c(4, 5), pty='s', mai=c(0.1, 0, 0.15, 0.1))
for (lab in 0:9) {
  samp <- train %>%
    filter(y == lab)
  for (i in 1:2) {
    img <- matrix(as.numeric(samp[i, -1]), 28, 28, byrow = TRUE)
    image(t(rotate(img))[,28:1], axes = FALSE, col = grey(seq(1, 0, length = 256)),xlim = c(1,0))
    box(lty = 'solid')
    title(main = lab, font.main=22)
  }
}
```




## C)
**Train a simple fully connected network with a single hidden layer to predict the digits. Do not forget to normalize the data similarly to what we saw with FMNIST in class.**


```{r Build simple model}

train_x<-array_reshape(train_x/255,c(nrow(train_x),784))
test_x<-array_reshape(test_x/255,c(nrow(test_x),784))

train_y <- to_categorical(train_y, num_classes = 10)
test_y <- to_categorical(test_y, num_classes = 10)

simple_keras <- keras_model_sequential()
simple_keras |>
    layer_dense(units = 128, activation = 'relu', input_shape = c(784)) |>
    layer_dropout(rate = 0.2) |>
    layer_dense(units = 10, activation = 'softmax')

summary(simple_keras)
```

```{r train simple model}

compile(
    simple_keras,
    loss = 'categorical_crossentropy',
    optimizer = optimizer_adam(),
    metrics = c('accuracy')
)

fit(
    simple_keras, train_x, train_y,
    epochs = 20, batch_size = 20,
    validation_data = list(test_x, test_y)
)

```

```{r Evaluate simple model}
evaluate(simple_keras, test_x, test_y)
```


## D)
**Experiment with different network architectures and settings (number of hidden layers, number of nodes, type and extent of regularization, etc.). Train at least 5 models. Explain what you have tried, what worked and what did not. Make sure that you use enough epochs so that the validation error starts flattening out - provide a plot about the training history.**




## E) 
**Choose a final model and evaluate it on the test set. How does test error compare to validation error?**


# Problem 2

## A)
**Show two images: one hot dog and one not hot dog. (Hint: You may use knitr::include_graphics() or install the imager package to easily accomplish this.)**

```{r Get pictures}
#Show hot dog
path<-"C:/CEU/DS1/ceu-ds1/data/train/"
hot_dog<-load.image(paste0(path,"hot_dog/7896.jpg"))
plot(hot_dog)

# Show not hot dog
not_hot_dog<-load.image(paste0(path,"not_hot_dog/4770.jpg"))
plot(not_hot_dog)
```

## B)
**What would be a good metric to evaluate such a prediction?**


## C)
**To be able to train a neural network for prediction, let’s first build data batch generator functions as we did in class for data augmentation (train_generator and valid_generator). **


```{r data batch generator}
batch_size <- 16

train_datagen <- image_data_generator(rescale = 1/255)

valid_datagen <- image_data_generator(rescale = 1/255)

train_generator <- flow_images_from_directory(
    directory="C:/CEU/DS1/ceu-ds1/data/train/",
    class_mode = "binary",
    generator = train_datagen,
    batch_size = batch_size,
    target_size = c(128, 128)
)

valid_generator <- flow_images_from_directory(
    directory="C:/CEU/DS1/ceu-ds1/data/test/",
    class_mode = "binary",
    generator = valid_datagen,
    batch_size = batch_size,
    target_size = c(128, 128)
)

```

## D)
**Build a simple convolutional neural network (CNN) to predict if an image is a hot dog or not. Evaluate your model on the test set. (Hint: Account for the fact that you work with colored images in the input_shape parameter: set the third dimension to 3 instead of 1.)**


```{r Build CNN for hot dogs}

cnn_model_w_augmentation <- keras_model_sequential()
cnn_model_w_augmentation |>
    layer_conv_2d(
        filters = 32,
        kernel_size = c(3, 3),
        activation = 'relu',
        input_shape = c(128, 128, 3)
    ) |>
    layer_max_pooling_2d(pool_size = c(2, 2)) |>
    layer_dropout(rate = 0.2) |>
    layer_flatten() |>
    layer_dense(units = 32, activation = 'relu') |>
    layer_dense(units = 1, activation = 'softmax')

compile(
    cnn_model_w_augmentation,
    loss = 'binary_crossentropy',
    optimizer = optimizer_adam(),
    metrics = c('accuracy')
)

fit(
    cnn_model_w_augmentation,
    train_generator,
    epochs = 20,
    steps_per_epoch = 10, #nrow(data_train_x) / batch_size,  # this does not make a difference here -- batch_size of the generator determines how training works
    validation_data = valid_generator,
    validation_steps = 10#nrow(data_valid_x) / batch_size
)
```

The train and test accuracy are both about 50% which means, that this binary classification model is not better than random :(. 



## E)
**Could data augmentation techniques help with achieving higher predictive accuracy? Try some augmentations that you think make sense and compare to your previous model.**

## optional
**Try to rely on some pre-built neural networks to aid prediction. Can you achieve a better performance using transfer learning for this problem?**






